---
title: "STA457 Assignment 1"
author: "Aofei Liu (1002291334) Yu-Hong Lin (1002945218)"
date: "October 27, 2018"
output:
  pdf_document: default
  word_document: default
---

Question 1. Derive the variance of the predictor $F_t$ given in Equation(1).

\textbf{Answer:}

$$
\begin{aligned}
\sigma^2_F &= var(\sum_{i=0}^{m-2} d_i \times X_{t-i})\\
\\
&= \sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times Cov(X_{t-i}, X_{t-j})\\
\\
&= \sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times \gamma_{(i-j)}\\
\\
&= \sum_{i=0}^{m-2} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-j)}) \bigg \}\\
\\
&= \sum_{i=0}^{r-1} d_i \bigg \{\sum_{j=0}^{r-1} d_j \gamma_{(i-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-j)} \bigg \} \\
\\
&+ \sum_{i=r}^{m-2} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-j)}\bigg \}\\
\\
&= \sum_{i=0}^{r-1}\sum_{j=0}^{r-1} d_i d_j \gamma_{(i-j)} + 2 \times \bigg \{ \sum_{i=0}^{r-1}\sum_{j=r}^{m-2} d_i d_j \gamma_{(i-j)} \bigg \} + \sum_{i=r}^{m-2}\sum_{j=r}^{m-2}d_id_j \gamma_{(i-j)}
\end{aligned}
$$

\newpage

Question 2. Derive the expectation of $F_t$.

\textbf{Answer:}

$$
\begin{aligned}
\mu_F &= E(\sum_{i=0}^{m-2} d_i \times X_{t-i})\\
\\
&= \sum_{i=0}^{m-2} d_i \times E(X_{t-i})\\
\\
&= \sum_{i=0}^{m-2} d_i \times  \mu_X\\
\\
&= \bigg \{ \sum_{i=0}^{r-1} d_i \times \mu_X \bigg \} + \bigg\{ \sum_{i=r}^{m-2} d_i \times \mu_X \bigg \} \\
\end{aligned}
$$

\newpage

Question 3. Derive the autocorrelation function at lag one for the predictor.

\textbf{Answer:}

$$
\begin{aligned}
\rho_F(1) &= corr(F_t, F_{t-1}) = \frac{\gamma_F(1)}{\gamma_F(0)} = \frac{\gamma_F(1)}{\sigma_F^2} \ \ \ \ \ \ \  Note: \gamma_F(0) = Var(F_t) = \sigma_F^2\\
\\
\gamma_F(1) &= cov(F_t, F_{t-1})\\
\\
&= cov(\sum_{i=0}^{m-2} d_i X_{t-i}, \sum_{j=0}^{m-2} d_j X_{t-1-j})\\
\\
&\approx E(\sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times X_{t-i} X_{t-1-j})\\
\\
&= \sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times cov(X_{t-i},X_{t-1-j})\\
\\
&= \sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times \gamma_{(i-1-j)}\\
\\
&= \sum_{i=0}^{m-2} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-1-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-1-j)} \bigg \}\\
\\
&= \sum_{i=0}^{r-1} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-1-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-1-j)} \bigg \} \\
\\
&+ \sum_{i=r}^{m-2} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-1-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-1-j)} \bigg \}
\end{aligned}
$$

Given the value/expression of $\sigma_F^2$ from answer to Question 1,

$$
\begin{aligned}
\rho_F(1) &= \frac{1}{\sigma_F^2} \times \bigg \{ \sum_{i=0}^{r-1} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-1-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-1-j)} \bigg \} \\
\\
&+ \sum_{i=r}^{m-2} d_i \bigg \{ \sum_{j=0}^{r-1} d_j \gamma_{(i-1-j)} + \sum_{j=r}^{m-2} d_j \gamma_{(i-1-j)} \bigg \} \bigg \} \\
&(where \ \sigma^2_F \ was \ derived \ in \ Question \ 1.)
\\
\end{aligned}
$$

\newpage

Question 4. Write a R function to calculate the expectation of the rule return for a given double MA trading rule (See Appendix) and the expected length of the holding period.

\textbf{Answer:}

Notes, we need to derive a formula for $corr(X_t, F_{t-1})$ to compute $E(R_t)$.

$$
\begin{aligned}
corr(X_t, F_{t-1}) &= \frac{cov(X_t, F_{t-1})}{\sqrt{var(X_t) \times var(F_{t-1})}} \\
\\
&=  \frac{cov(X_t,  \sum_{j=0}^{m-2} d_j X_{t-1-j})}{\sqrt{var(X_t) \times var(F_{t-1})}} \\
\\
&= \frac{\sum _{i=0}^{m-2} \gamma_{(i+1)}}{\sqrt{\gamma_0 \{\sum_{i=0}^{m-2} \sum_{j=0}^{m-2} d_i \times d_j \times \gamma_{(i-j)} \}}}
\end{aligned}
$$


```{r eval = TRUE, echo =TRUE}
X_func <- function(price){
  x_vec <- c()
  for (i in 1: length(price)-1){
    x_vec[i] <- log(price[i] / price[i+1])
  }
  return(x_vec)
}

intergrand <- function(x){
  res <- (sqrt(2 * pi))^(-1) * exp(-x^2 / 2)
  return(res)
}

phi_func <- function(h){
  res <- integrate(intergrand, lower = -Inf, upper = h)$value
  return(res)
}

d_func <- function(m,r){
  res <- c()
  for (i in 1: (r)){
    res[i] <-(m-r) * (i)
  }
  for (i in (r+1):(m-1)){
    res[i] <- r * (m-i)
  }
  return(res)
}

sigma_f_func <- function(m, r, gamma_x){
  d <- d_func(m,r)
  res <- 0
  for (i in 1: (m-1)){
    for (j in 1: (m-1)){
      res <- res + d[i] * d[j] * gamma_x[abs(i-j)+1]
    }
  }
  res <- sqrt(res)
  return(res)
}

mu_f_func <- function(m,r,mu_x){
  d <- d_func(m,r)
  res <- sum(d) * mu_x
  return(res)
}

corr_xf_func <- function(m,r,gamma_x){
  d <- d_func(m,r)
  a <- sigma_f_func(m,r,gamma_x)
  res <- sum(d * gamma_x[2:m]) / (sqrt(gamma_x[1]) * a)
  return(res)
}

expectation <- function(s,m,r){
  GSPC_rev <- as.data.frame(s)
  GSPC_rev <- GSPC_rev[dim(GSPC_rev)[1]:1,]

  x <- X_func(GSPC_rev$GSPC.Adjusted)
  mu_x <- mean(x)
  sigma_x <- sd(x)
  gamma_x <- as.numeric(acf(x, type = "covariance", plot=FALSE, lag.max = length(x))$acf)

  d <- d_func(m,r)
  mu_f <- mu_f_func(m,r,mu_x)
  sigma_f <- sigma_f_func(m,r,gamma_x)
  frac_f <- mu_f / sigma_f
  phi <- phi_func(-frac_f)
  corr <- corr_xf_func(m,r,gamma_x)
  ert <- sqrt(2/pi) * sigma_x * corr * exp(-0.5 * (frac_f ^ 2)) + mu_x * (1 - 2*phi)

  h <- 0
  d <- d_func(m,r)
  for (i in 1: (m-1)){
    for (j in 1: (m-1)){
      h <- h + d[i] * d[j] * gamma_x[abs(i-j-1)+1]
    }
  }
  h <- h / (sigma_f ^2)
  h <- pi / acos(h)

  return(list(ert = ert, h = h))
}

```

\newpage

Question 5. Use a R function to download daily, weekly S&P500 index from Oct/01/2009 to Sep/30/2018 from yahoo finance.

\textbf{Answer:}

```{r eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE}
library("quantmod")
GSPC_daily <- getSymbols(Symbols = "^GSPC", src = "yahoo", auto.assign = FALSE,
                         from = "2009-10-01", to = "2018-09-30")


GSPC_weekly <- getSymbols(Symbols = "^GSPC", src = "yahoo", auto.assign = FALSE,
                          from = "2009-10-01", to = "2018-09-30", periodicity = "weekly")

cat("daily S&P500 index", "\n")

summary(GSPC_daily)

cat("\n","\n")
cat("weekly S&P500 index", "\n")

summary(GSPC_weekly)
```


\newpage

Question 6. Write a R function to choose the optimal daily and weekly double MA trading rules (that maximize the expected rule returns) for S&P500 index.

\textbf{Answer:}

```{r eval = TRUE, echo = FALSE}
pair <- function(m_max){
  #m-2 >= r
  m <- seq(3,m_max,by=1)
  p <- list()
  for (i in 1:length(m)){
    a <- seq(1,m[i]-2,by=1)
    p[[i]] <- expand.grid(m[i],a)
  }
  res <- do.call(rbind, p)
  colnames(res) <- c("m","r")
  return(res)
}

optimal_rt <- function(s,m_max){
  pair_mr <- pair(m_max)
  expected_return_vec <- c()
  for (i in 1: nrow(pair_mr)){
    m_iter <- as.numeric(pair_mr[i,])[1]
    r_iter <- as.numeric(pair_mr[i,])[2]
    expected_return_vec[i] <- expectation(s,m_iter,r_iter)[1]
  }
  max_mr <- pair_mr[which.max(expected_return_vec),]
  return(max_mr)
}

m_max_day <- 50
mr <- optimal_rt(GSPC_daily, m_max_day)
m_max_week <- 52
mr_w <- optimal_rt(GSPC_weekly, m_max_week)
```

```{r eval = TRUE, echo = FALSE}
cat("daily S&P500 index","\n")
cat("mr pair", "(", as.numeric(mr[1]), ",", as.numeric(mr[2]), ")", "has highest E(Rt)", "\n", "\n")

cat("weekly S&P500 index","\n")
cat("mr pair", "(", as.numeric(mr_w[1]), ",", as.numeric(mr_w[2]), ")", "has highest E(Rt)", "\n")
```

Therefore, optimal daily double MA trading rules are (44,40) and optimal weekly double MA trading rules are (52,50).

\newpage

Question 7. Write a R function to calculate the in-sample trading statistics (cumulative return and holding time) of your choice and compare them with your theoretical results.

\textbf{Answer:}

```{r eval= TRUE, echo = FALSE}
trading_statistics <- function(s,m,r){
  GSPC_rev <- as.data.frame(s)
  GSPC_rev <- GSPC_rev[dim(GSPC_rev)[1]:1,]
  price <- GSPC_rev$GSPC.Adjusted
  x <- X_func(price)
  n <- length(price) - (m-1)
  d <- d_func(n+1,r)
  f <- c()
  for (i in 1:(n)){
    f[i] <- sum(d[i:n] * x[(i+1):(n+1)])
  }
  b <- ifelse(f > 0, 1, -1)
  sum_r_t <- sum(b * x[1:n])
  res <- sum_r_t / n
  count <-  sum(diff(sign(b)) != 0)
  count <- n/count
  return(list(crt=res, avht = count))
}

mr <- c(250,248)
expect_return_d <- expectation(GSPC_daily, as.numeric(mr)[1], as.numeric(mr)[2])$ert
expect_holding_time <- expectation(GSPC_daily, as.numeric(mr)[1], as.numeric(mr)[2])$h
cumulative_return_d <-  trading_statistics(GSPC_daily, as.numeric(mr)[1], as.numeric(mr)[2])$crt
average_holding <-  trading_statistics(GSPC_daily, as.numeric(mr)[1], as.numeric(mr)[2])$avht

cat("daily S&P500 index","\n")
cat("theoretical cumulative return:" ,expect_return_d, "\n")
cat("sample cumulative return:", cumulative_return_d,"\n","\n")

cat("theoretical holding time:", expect_holding_time,"\n")
cat("sample holding time:", average_holding,"\n","\n")
```

```{r eval= TRUE, echo = FALSE, warning = FALSE}

expect_returnw <- expectation(GSPC_weekly, as.numeric(mr_w)[1], as.numeric(mr_w)[2])$ert
expect_holding_time_w <- expectation(GSPC_weekly, as.numeric(mr_w)[1], as.numeric(mr_w)[2])$h
cumulative_return_w <- trading_statistics(GSPC_weekly, as.numeric(mr_w)[1], as.numeric(mr_w)[2])$crt
average_holding_w <- trading_statistics(GSPC_weekly, as.numeric(mr_w)[1], as.numeric(mr_w)[2])$avht


cat("weekly S&P500 index","\n")
cat("theoretical cumulative return:" ,expect_returnw, "\n")
cat("sample cumulative return:", cumulative_return_w,"\n","\n")

cat("theoretical holding time:", expect_holding_time_w,"\n")
cat("sample holding time:", average_holding_w,"\n")
```

For daily S&P500 index, theoretical cumulative return (0.0003198645) is similiar to average cumulative return (0.000256746). Theoretical holding time (23.07712) is much smaller than average holding time (61.09091).

For weekly S&P500 index, theoretical cumulative return (0.001520497) is similiar to average cumulative return (0.001612475). Theoretical holding time (11.31526) is much smaller than average holding time (69.83333).
